---
title: "Limma Tutorial"
author: Lindsay Rutter
output:
  packagedocs::package_docs:
  toc: true
toc_collapse: true
vignette: >
  %\VignetteEngine{packagedocs::redirect}
---
  
  <meta http-equiv="content-type" content="text/html;charset=utf-8" />
  
  ```{r global_options, include=FALSE}
# R output pre blocks are styled by default to indicate output
knitr::opts_chunk$set(
  comment = NA,
  cache = TRUE,
  fig.height = 8,
  fig.width = 10
)
```
Try new paper (https://www.bioconductor.org/help/workflows/RNAseq123/)

This script creates the following three datasets:
"topGenes_100limma.Rds"
"genePval_100limma.Rds"
"data_100limma.Rds"

```{r}
library(limma)
library(Glimma)
library(GGally)
library(ggplot2)

thisPath <- "/Users/lindz/bigPint"
beeCounts <- read.delim(file=paste0(thisPath, "/AllLaneCount.txt"), row.names=1, stringsAsFactors = FALSE)
colnames(beeCounts) <- c("NC.1", "NC.2", "NR.1", "VR.1", "NS.1", "VP.1", "NS.2", "VR.2", "NP.1", "VP.2", "VC.1", "NP.2", "VP.3", "NP.3", "VS.1", "VS.2", "VC.2", "NC.3", "VP.4", "NC.4", "NR.2", "VC.3", "VC.4", "NP.4", "VR.3", "NC.5", "VS.3", "NP.5", "VC.5", "VS.4", "NS.3", "VS.5", "VP.5", "NR.3", "NR.4", "VC.6", "NS.4", "NC.6", "NP.6", "VR.4", "NR.5", "NR.6", "NS.5", "VP.6", "NS.6", "VR.5", "VR.6", "VS.6")
beeCounts <- beeCounts[ , order(names(beeCounts))]
x <- DGEList(counts=beeCounts)

#samplenames <- substring(colnames(x), 12, nchar(colnames(x)))
#colnames(x) <- samplenames
group <- as.factor(colnames(x))
group <- factor(c(rep("NC",6), rep("NP",6), rep("NR",6), rep("NS",6), rep("VC",6), rep("VP",6), rep("VR",6), rep("VS",6)))
x$samples$group <- group
lane <- as.factor(c("L12","L12","L12","L12","L34","L34","L12","L12","L12","L12","L34","L34","L12","L12","L34","L34","L34","L34","L12","L12","L34","L34","L34","L34","L12","L12","L12","L12","L34","L34","L12","L12","L12","L12","L34","L34","L12","L12","L34","L34","L34","L34","L12","L12","L34","L34","L34","L34"))
x$samples$lane <- lane
```

Transform and remove low counts.
```{r}
cpm <- cpm(x)
lcpm <- cpm(x, log=TRUE)
keep.exprs <- rowSums(cpm>1)>=8 # tried filtering up to 24 and not much difference
x <- x[keep.exprs,, keep.lib.sizes=FALSE] # 15,314 to 10,626
x <- calcNormFactors(x, method = "TMM")
```

Create design matrix. There are many ways to setup a design matrix. Here, we removed the intercept from group (the first factor), but kept the intercept from lane. This allows us to do contrasts with group more easily.

```{r}
design <- model.matrix(~0+group+lane)
colnames(design) <- gsub("group", "", colnames(design))

contr.matrix <- makeContrasts(
   NCvsNP = NC-NP,
   NCvsNR = NC-NR,
   NCvsNS = NC-NS,
   NCvsVC = NC-VC,
   NCvsVP = NC-VP,
   NCvsVR = NC-VR,
   NCvsVS = NC-VS,
   NPvsNR = NP-NR,
   NPvsNS = NP-NS,
   NPvsVC = NP-VC,
   NPvsVP = NP-VP,
   NPvsVR = NP-VR,
   NPvsVS = NP-VS,
   NRvsNS = NR-NS,
   NRvsVC = NR-VC,
   NRvsVP = NR-VP,
   NRvsVR = NR-VR,
   NRvsVS = NR-VS,
   NSvsVC = NS-VC,
   NSvsVP = NS-VP,
   NSvsVR = NS-VR,
   NSvsVS = NS-VS,
   VCvsVP = VC-VP,
   VCvsVR = VC-VR,
   VCvsVS = VC-VS,
   VPvsVR = VP-VR,
   VPvsVS = VP-VS,
   VRvsVS = VR-VS,
   levels = colnames(design))
```

```{r}
v <- voom(x, design)
vfit <- lmFit(v, design)
vfit <- contrasts.fit(vfit, contrasts=contr.matrix)
efit <- eBayes(vfit)
```

For a quick look at differential expression levels.

```{r}
summary(decideTests(efit))
```

```{r}
pairNames <- colnames(contr.matrix)
topGenes <- list()
genePval <- list()
for (i in 1:length(pairNames)) {
  temp <- topTreat(efit, coef=i, n=Inf)
  temp2 <- temp[order(temp[,5]),]
  sigRows <- 1:100
  temp3 <- temp2[sigRows,]
  setDT(temp3, keep.rownames = TRUE)[]
  colnames(temp3)[1] = "ID"
  colnames(temp3)[5] = "pVal" # can't have dots in name
  colnames(temp3)[6] = "adjPVal" # can't have dots in name
  topGenes[[pairNames[i]]] <- as.data.frame(temp3)
}
```

Save the data

```{r}
saveRDS(topGenes, file=paste0(getwd(), "/topGenes_100limma.Rds"))
saveRDS(x, file=paste0(getwd(), "/data_100limma.Rds"))
```
