---
title: "Honey Bees - edgeR (check p-values before blocking)"
author: Lindsay Rutter
output:
  packagedocs::package_docs:
    toc: true
    toc_collapse: true
vignette: >
  %\VignetteEngine{packagedocs::redirect}
---

<meta http-equiv="content-type" content="text/html;charset=utf-8" />

```{r global_options, include=FALSE}
# R output pre blocks are styled by default to indicate output
knitr::opts_chunk$set(
  comment = NA,
  cache = TRUE,
  fig.height = 8,
  fig.width = 10
)
```

```{r}
library(edgeR)
library(ggplot2)
library(GGally)
library(EDASeq)
library(utils)

thisPath <- "/Users/lindz/BeeVirusDiet/edgeR"

beeCounts <-read.delim(file="../AllLaneCount.txt",row.names=1,stringsAsFactors = FALSE)

colnames(beeCounts) <- c("NC.1", "NC.2", "NR.1", "VR.1", "NS.1", "VP.1", "NS.2", "VR.2", "NP.1", "VP.2", "VC.1", "NP.2", "VP.3", "NP.3", "VS.1", "VS.2", "VC.2", "NC.3", "VP.4", "NC.4", "NR.2", "VC.3", "VC.4", "NP.4", "VR.3", "NC.5", "VS.3", "NP.5", "VC.5", "VS.4", "NS.3", "VS.5", "VP.5", "NR.3", "NR.4", "VC.6", "NS.4", "NC.6", "NP.6", "VR.4", "NR.5", "NR.6", "NS.5", "VP.6", "NS.6", "VR.5", "VR.6", "VS.6")

beeCounts <- beeCounts[ , order(names(beeCounts))]

y <- DGEList(counts=beeCounts)
```

edgeR vignette states that a gene is required to have a count of 5-10 in a library to be considered expressed in that library. Here minLib is equal to `r min(y$samples$lib.size)`. A CPM of 3 corresponds to a count of ~9 in the minimum number of samples in a group (24). So, I keep only rows that have that. This reduces the number of genes from 15,314 to 8,581.

```{r}
minLib <- min(y$samples$lib.size)
keep <- rowSums(cpm(y)>3) >= 24
# Number of genes 15,314--> 8,581
y <- y[keep, , keep.lib.sizes=FALSE]
```

Next, I used edgeR normalization.
```{r}
y <- calcNormFactors(y)
```

I can create boxplot and RLE boxplot.

```{r, echo=FALSE, eval=FALSE}
ggparcoord(data.frame(y[[1]]), columns=1:48, alphaLines=0, boxplot=TRUE, scale="globalminmax") + coord_flip()

ggparcoord(data.frame(log(y[[1]]/colMeans(y[[1]]))), columns=1:48, alphaLines=0, boxplot=TRUE, scale="globalminmax") + coord_flip()
```

I can create MDS plots of interest.

```{r, echo=FALSE, eval=FALSE}
allGroups <- c(rep("NC",6), rep("NP",6), rep("NR",6), rep("NS",6), rep("VC",6), rep("VP",6), rep("VR",6), rep("VS",6))
y$samples$group <- allGroups

plotMDS(y, col = c("red","deeppink","darkorange","gold","green2", "green4","blue", "purple")[factor(allGroups)], cex=0.6)

plotMDS(y, col = c("blue","blue","blue","blue","red","red","red","red")[factor(allGroups)], cex=0.6)

plotMDS(y, col = c("white","blue","white","red","white","white","white","white")[factor(allGroups)], cex=0.6)

plotMDS(y, col = c("white","white","white","white","white","blue","white","red")[factor(allGroups)], cex=0.6)

plotMDS(y, col = c("white","blue","white","white","white","white","white","red")[factor(allGroups)], cex=0.6)
```

Copying Section (3.4.2 in edgeR vignette)

```{r}
Group = factor(c(rep("NC",6), rep("NP",6), rep("NR",6), rep("NS",6), rep("VC",6), rep("VP",6), rep("VR",6), rep("VS",6)))
design <- model.matrix(~0+Group, data=y$samples)
colnames(design) <- levels(Group)
y <- estimateDisp(y, design)
plotBCV(y)
```

```{r}
fit <- glmFit(y, design)

myContrasts <- makeContrasts(
  VvsN = (VC+VS+VP+VR)-(NC+NS+NP+NR),
  NSvsNP = NS-NP,
  VSvsVP = VS-VP,
  NPvsVS = NP-VS,
  NSNPvVSVP = (NS-NP)-(VS-VP),
levels=design)
```

Below we save the DEGs from all pairwise combinations of treatment groups.

```{r}
allPairs = data.frame(Treatment1 = factor(), Treatment2 = factor(), NumberDEG = numeric(), FirstLarger = numeric(), SecondLarger = numeric())

for (i in 1:(ncol(fit)-1)){
  for (j in (i+1):ncol(fit)){
    contrast=rep(0,ncol(fit))
    contrast[i]=1
    contrast[j]=-1
    lrt <- glmLRT(fit, contrast=contrast)
    lrt <- topTags(lrt, n = nrow(y[[1]]))[[1]]
    
    pVals <- data.frame(x = lrt$PValue)
    ggplot(lrt$Pvalue) + geom_histogram()
    ggplot(pVals, aes(x=x)) + geom_histogram()
    
    png(filename=paste0(thisPath, "/pvalPlots/", colnames(fit)[i], colnames(fit)[j],".Rds"))
    ggplot(pVals, aes(x=x)) + geom_histogram() + xlab("p-value") + ylab("Gene count") + ggtitle(paste0(colnames(fit)[1], " versus ", colnames(fit)[2]))
    dev.off()
    
    saveRDS(lrt, file=paste0(thisPath, "/DEGPairs/", colnames(fit)[i], colnames(fit)[j],".Rds"))
    
    lrt <- lrt[which(lrt$FDR<0.05),]
    lrtLength <- nrow(lrt)
    allPairs = rbind(allPairs, data.frame(Treatment1 = factor(colnames(fit)[i]), Treatment2 = factor(colnames(fit)[j]), NumberDEG = lrtLength))
  }
}
allPairs <- allPairs[order(allPairs$NumberDEG),]
saveRDS(allPairs, file=paste0(thisPath, "/DEGPairs/allPairs.Rds"))
```

```{r}
allPairs
```